---
title: "Roads in Sarasota"
author: "Angela Kothe"
date: "5/2/2022"
output: html_document
---

I made a couple maps of roads in Sarasota using the ggmap and viridis packages. The underlying data was available on the Explore Sarasota County website, as was the coast and road shapefiles.

I first made a map of streets, with colors indicating the speed limit.

```{r, width = 25, height = 20, echo = FALSE, show_col_types = FALSE, fig.keep = "last", message = FALSE, results = "hide", warning = FALSE, fig.height = 20, fig.width = 25}
library(maptools)
library(rgeos)
library(tidyverse)
library(rgdal)
library(ggthemes)
library(socviz)
library(scales)
library(cowplot)
library(ggmap)
library(dplyr)
library(showtext)
library(sf)
library(extrafont)
library(haven)
library(here)
library(viridis)

#streets and roads shapefile; data retrieved from shape
s <- readOGR(here("shapes", "streets", "Street.shp"))
s <- spTransform(s, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
s.df <- fortify(s, region = "zipr")

c <- readOGR(here("shapes", "coast", "CoastalRegulatoryLine.shp"))
c <- spTransform(c, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
c.df <- fortify(c)

data <-s@data
data$objectid <- as.character(data$objectid)
data <- data %>%
          rename(id = objectid)

#merge data with shapefile
data <- left_join(s.df, data, by = "id", sort = FALSE)

#plot
font_add_google(name = "Lato", family = "lato") 
showtext_auto()

ggplot() +
    geom_path(data = c.df,
              aes(long, lat, group = group), 
              colour = "skyblue", 
              size = .2) +
    geom_path(data = data, aes(x = long, y = lat, group = group,
                               color = speedlimit),
              size = .3, 
              alpha = 1) +
    coord_sf(xlim = c(-82.7, -82.4), 
             ylim = c(27.2, 27.4),
             expand = FALSE) +
    scale_fill_viridis_c(alpha = 1,
                         begin = 0,
                         end = 1,
                         direction = -1,
                         option = "H",
                         values = NULL,
                         space = "Lab",
                         na.value = "grey50",
                         guide = "colourbar",
                         aesthetics = "colour") +
    theme_map() +
    theme(legend.position = "top", 
          legend.justification = "center",
          legend.title = element_blank(),
          legend.text = element_text(color = "white", 
                                     size = 28, 
                                     family = "lato"),
          legend.key.height = unit(1, "cm"),
          legend.key.width = unit(3, "cm"),
          plot.title = element_text(size = 40, 
                                    family = "lato", 
                                    face = "bold", 
                                    hjust = .5,
                                    color = "white"),
          plot.background = element_rect(fill = "black", 
                                         color = "black"),
          plot.subtitle = element_text(family = "lato", 
                                       size = 16, 
                                       hjust = .5, 
                                       color = "white")) +
    labs(title = "STREETS IN SARASOTA", 
         subtitle = "Speed Limit")

```

I also made a map of the eight most common road types in Sarasota and color coded the data.

```{r, width = 25, height = 20, echo = FALSE, show_col_types = FALSE, fig.keep = "last", message = FALSE, results = "hide", warning = FALSE, fig.height = 20, fig.width = 25}

#color palette
colors <- c("skyblue", "hotpink", "green", "red", "orange", "purple", 
            "yellow", "pink", "white")

data <- read_dta(here("data", "streets.dta"))
data <- data %>%
          rename(id = objectid)

#merge data with shapefile
data$id <- as.character(data$id)
data <- left_join(s.df, data, by = "id", sort = FALSE)
data$type[is.na(data$type)] = "Other"

#plot font
font_add_google(name = "Lato", family = "lato") 
showtext_auto()

#plot
ggplot() +
      geom_path(data = data, 
                aes(x = long, y = lat, group = group, color = type),
                size = .3, alpha = 1) +
      coord_sf(xlim = c(-82.7, -82.4), 
               ylim = c(27.2, 27.4),
               expand = FALSE) +
      scale_color_manual(values = colors) + 
      theme_map() +
      theme(legend.position = "top", 
            legend.justification = "center",
            legend.title = element_blank(),
            legend.text = element_text(color = "white", 
                                       size = 28, 
                                       family = "lato"),
            legend.key.height = unit(1, "cm"),
            legend.key.width = unit(1, "cm"),
            plot.title = element_text(size = 40, 
                                      family = "lato", 
                                      face = "bold", 
                                      hjust = .5,
                                      color = "white"),
            plot.background = element_rect(fill = "black", 
                                           color = "black"),
            plot.subtitle = element_text(family = "lato", 
                                         size = 16, 
                                         hjust = .5, 
                                         color = "white")) +
      labs(title = "STREETS IN SARASOTA", 
             subtitle = "Street Type")

```

The map illuminates the loose application of road classifications. A road is
any pathway that connects two given points, and the most general kind of path designated for cars. An avenue is a road that is perpendicular to a street with trees or buildings on the sides. Similarly, a street is a road that runs perpendicular to an avenue and has buildings along its side. A court is a road that ends in a cul-de-sac. A circle is a road that is a closed loop. If you look at the map, many roads subvert their proper definition.




