---
title: "Trees in Sarasotan Zip Codes"
author: "Angela Kothe"
date: "5/2/2022"
output: html_document
---

I made maps of trees in Sarasota by zip code. I used data accessible on the Explore Sarasota County database. I downloaded street, zip code boundaries, and coastal shapefiles; as well as a dataset of trees planted. I coded the map to randomly distribute the trees according to the zip code they were planted in. This allowed me to make dot-density maps, so that I could display thousands of data points clearly. The code to randomly distribute the dots was borrowed from the Culture of Insight blog.

I started by making a map of trees in Sarasota by height, with one point representing twenty trees. There are zip codes seemingly without data, likely as a result of the random distribution or limitations of the dataset.

```{r, width = 25, height = 20, echo = FALSE, show_col_types = FALSE, fig.keep = "last", message = FALSE, results = "hide", warning = FALSE, fig.height = 20, fig.width = 25}

library(maptools)
library(rgeos)
library(tidyverse)
library(rgdal)
library(ggthemes)
library(socviz)
library(scales)
library(cowplot)
library(ggmap)
library(dplyr)
library(showtext)
library(sf)
library(extrafont)
library(haven)
library(here)

#wd <- here("Documents", "Github", "pbl")

#data, tree height; zip code and roads shapefile
data <- read_dta(here("data", "treeheight.dta"))
data$zip_code <- as.character(data$zipcode)

map <- readOGR(here("shapes", "zips", "ZipCode.shp"))
map <- spTransform(map, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
map.df <- fortify(map, region = "zip_code")

s <- readOGR(here("shapes", "streets", "Street.shp"))
s <- spTransform(s, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
s.df <- fortify(s, region = "zipr")

c <- readOGR(here("shapes", "coast", "CoastalRegulatoryLine.shp"))
c <- spTransform(c, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
c.df <- fortify(c)

#palette
colors <- c("#fd7c84", "#025b0e", "#9db802", "#000080", "#80b1d3", 
            "mediumvioletred", "mediumpurple1")

#plot
mdata <- left_join(map@data, data, by = "zip_code", sort = FALSE)
mdata[is.na(mdata)] <- 0

num.dots <- select(mdata, less10:h6070) / 20
num.dots <- round(num.dots, 0)

sp.dfs <- lapply(names(num.dots), function(x) {
  dotsInPolys(map, as.integer(num.dots[, x]), f="random")
})


dfs <- lapply(sp.dfs, function(x) {
  data.frame(coordinates(x)[,1:2])
})

type <- c("Less than Ten", "Ten to Twenty", "Twenty to Thirty", 
          "Thirty to Forty", "Forty to Fifty", "Fifty to Sixty", 
          "Greater than Sixty")

for (i in 1:length(type)) {
  dfs[[i]]$type <- type[i]
}

dots.final <- bind_rows(dfs)
dots.final$type <- factor(dots.final$type, levels = type)

font_add_google(name = "Lato", family = "lato") 
showtext_auto()

height <- ggplot(s.df) +
      geom_path(data = c, aes(long, lat, group = group), 
                colour = "skyblue", 
                size = 1) +
      geom_path(aes(long, lat, group = group), 
                colour = "darkgrey", 
                size = .3) +
      coord_sf(xlim = c(-82.7, -82.4), 
               ylim = c(27.2, 27.4),
               expand = FALSE) +
      geom_point(data = dots.final, 
                 aes(x, y, colour = type), 
                 size = 2) +
      scale_colour_manual(values = colors) +
      theme_map() +
      theme(legend.position = "top", 
            legend.justification = "center",
            legend.title = element_blank(),
            legend.text = element_text(color = "black", 
                                       size = 28, 
                                       family = "lato"),
            legend.key.height = unit(1, "cm"),
            legend.key.width = unit(1, "cm"),
            plot.title = element_text(size = 40, 
                                      family = "lato", 
                                      face = "bold", 
                                      hjust = .5,
                                      color = "black"),
            plot.background = element_rect(fill = "white", 
                                           color = NA),
            plot.subtitle = element_text(family = "lato", 
                                         size = 16, 
                                         hjust = .5, 
                                         color = "black")) +
      labs(title = "TREES IN SARASOTA", 
           subtitle = "Height in Feet \nOne Point is Twenty Trees"); height
```

I then made a map of trees according to the decade they were planted, with one point signifying ten trees.

```{r, width = 25, height = 20, echo = FALSE, show_col_types = FALSE, fig.keep = "last", message = FALSE, results = "hide", warning = FALSE, fig.height = 20, fig.width = 25}

#data, tree height; zip code and roads shapefile
data <- read_dta(here("data", "treedate.dta"))
data$zip_code <- as.character(data$zipcode)

#palette
colors <- c("mediumvioletred", "tan2", "#9db802", "mediumpurple1", "#bebada")

#plot
data <- subset(data, select = -c(d2010))

mdata <- full_join(map@data, data, by = "zip_code", sort = FALSE)
mdata[is.na(mdata)] <- 0

num.dots <- select(mdata, d1970:d2000) / 10

num.dots <- round(num.dots, 0)

sp.dfs <- lapply(names(num.dots), function(x) {
  dotsInPolys(map, as.integer(num.dots[, x]), f = "random")
})

dfs <- lapply(sp.dfs, function(x) {
  data.frame(coordinates(x)[,1:2])
})


year <- c("d1970", "d1980", "d1990", "d2000")
for (i in 1:length(year)) {
  dfs[[i]]$year <- year[i]
}

dots.final <- bind_rows(dfs)
dots.final$type <- factor(dots.final$year, levels = year)

font_add_google(name = "Lato", family = "lato") 
showtext_auto()

ggplot(s.df) +
  geom_path(data = c, aes(long, lat, group = group), 
            colour = "skyblue", 
            size = 1) +
  geom_path(aes(long, lat, group = group), 
            colour = "darkgrey", 
            size = .3) +
  coord_sf(xlim = c(-82.7, -82.4), 
           ylim = c(27.2, 27.4),
           expand = FALSE) +
  geom_point(data = dots.final, 
             aes(x, y, colour = type), 
             size = 2) +
  scale_colour_manual(values = colors,
                      labels = c("1970", "1980", "1990", "2000")) +
  geom_point(aes(x = -82.50446, 
                 y = 27.29673), 
             size = 2,
             colour = "mediumvioletred") +
  theme_map() +
  theme(legend.position = "top", 
        legend.justification = "center",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", 
                                   size = 28, 
                                   family = "lato"),
        legend.key.height = unit(1, "cm"),
        legend.key.width = unit(1, "cm"),
        plot.title = element_text(size = 40, 
                                  family = "lato", 
                                  face = "bold", 
                                  hjust = .5,
                                  color = "black"),
        plot.background = element_rect(fill = "white", 
                                       color = NA),
        plot.subtitle = element_text(family = "lato", 
                                     size = 16, 
                                     hjust = .5, 
                                     color = "black")) +
  labs(title = "TREES IN SARASOTA", 
       subtitle = "Decade Planted \nOne Point is Ten Trees")

```

I also made a map of the six most common tree species in the city, with one point signifying twenty trees.

```{r, width = 25, height = 20, echo = FALSE, show_col_types = FALSE, fig.keep = "last", message = FALSE, results = "hide", warning = FALSE, fig.height = 20, fig.width = 25}

#data, tree height; zip code and roads shapefile
data <- read_dta(here("data", "treetype.dta"))
data$zip_code <- as.character(data$zipcode)

#palette
colors <- c("#fd7c84", "#025b0e", "#9db802", "#000080", "#80b1d3", 
            "mediumvioletred", "mediumpurple1")

#plot
mdata <- left_join(map@data, data, by = "zip_code", sort = FALSE)
mdata[is.na(mdata)] <- 0

num.dots <- select(mdata, other:braz) / 20

sp.dfs <- lapply(names(num.dots), function(x) {
  dotsInPolys(map, as.integer(num.dots[, x]), f="random")
})

dfs <- lapply(sp.dfs, function(x) {
  data.frame(coordinates(x)[,1:2])
})

type <- c("other", "oak", "palm", "pine", "holly", "crape", "braz")
for (i in 1:length(type)) {
  dfs[[i]]$type <- type[i]
}

dots.final <- bind_rows(dfs)
dots.final$type <- factor(dots.final$type, levels = type)

font_add_google(name = "Lato", family = "lato") 
showtext_auto()

ggplot(s.df) +
  geom_path(data = c, aes(long, lat, group = group), 
            colour = "skyblue", 
            size = 1) +
  geom_path(aes(long, lat, group = group), 
            colour = "darkgrey", 
            size = .3) +
  coord_sf(xlim = c(-82.7, -82.4), 
           ylim = c(27.2, 27.4),
           expand = FALSE) +
  geom_point(data = dots.final, 
             aes(x, y, colour = type), 
             size = 2) +
  scale_colour_manual(values = colors) +
  theme_map() +
  theme(legend.position = "top", 
        legend.justification = "center",
        legend.title = element_blank(),
        legend.text = element_text(color = "black", 
                                   size = 28, 
                                   family = "lato"),
        legend.key.height = unit(1, "cm"),
        legend.key.width = unit(1, "cm"),
        plot.title = element_text(size = 40, 
                                  family = "lato", 
                                  face = "bold", 
                                  hjust = .5,
                                  color = "black"),
        plot.background = element_rect(fill = "white", 
                                       color = NA),
        plot.subtitle = element_text(family = "lato", 
                                     size = 16, 
                                     hjust = .5, 
                                     color = "black")) +
  labs(title = "TREES IN SARASOTA", 
       subtitle = "Species Planted \nOne Point is Twenty Trees")

```

